---
title: "SE Seminar"
author: "David Issa Mattos"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, cache = T)
```

# Packages

```{r message=F}
library(bpcs)
library(tidyverse)
library(knitr)
library(kableExtra)
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)
set.seed(99)
```

# Gorillas using touchscreen

> Huskisson, S.M., Jacobson, S.L., Egelkamp, C.L. et al. Using a Touchscreen Paradigm to Evaluate Food Preferences and Response to Novel Photographic Stimuli of Food in Three Primate Species (Gorilla gorilla gorilla, Pan troglodytes, and Macaca fuscata). Int J Primatol 41, 5--23 (2020). <https://doi.org/10.1007/s10764-020-00131-0>

Let's import the data and select only the gorillas

```{r message=F}
d <- readr::read_csv('data/touchscreen.csv') %>% 
  dplyr::filter(species_type=='Gorilla') %>% 
  dplyr::select(species_type, SubjectCode, Trial, image1, image2, image_chosen)
```

Let's create a column that says which image was selected 1 for image2 0 for image 1

```{r}
d <- d %>% 
  dplyr::mutate(y =ifelse(image_chosen==image1, 0, 1)) %>% 
  select(-image_chosen)
```

Let's also do some recoding to facilitate interpreting

```{r}
d$image1 <- dplyr::recode(d$image1, "Ca" = 'Carrot')
d$image1 <- dplyr::recode(d$image1, "Cu" = 'Cucumber')
d$image1 <- dplyr::recode(d$image1, "Tu" = 'Turnip')
d$image1 <- dplyr::recode(d$image1, "Gr" = 'Grape')
d$image1 <- dplyr::recode(d$image1, "To" = 'Tomato')
d$image1 <- dplyr::recode(d$image1, "Ap" = 'Apple')

d$image2 <- dplyr::recode(d$image2, "Ca" = 'Carrot')
d$image2 <- dplyr::recode(d$image2, "Cu" = 'Cucumber')
d$image2 <- dplyr::recode(d$image2, "Tu" = 'Turnip')
d$image2 <- dplyr::recode(d$image2, "Gr" = 'Grape')
d$image2 <- dplyr::recode(d$image2, "To" = 'Tomato')
d$image2 <- dplyr::recode(d$image2, "Ap" = 'Apple')
```

Sample of the data

```{r}
d %>% 
  dplyr::sample_n(size=10) %>% 
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "100%")
```

Now we can run a Bradley-Terry model with the bpcs package

```{r eval=F}
m1_gor <-
  bpc(
    d,
    player0 = 'image1',
    player1 = 'image2',
    result_column = 'y',
    model_type = 'bt',
    priors = list(prior_lambda_std = 0.5),
    iter = 3000
  )
save_bpc_model(m1_gor, 'm1_gor', 'fittedmodels')
```

```{r echo=F}
m1_gor <- load_bpc_model('fittedmodels/m1_gor.RDS')
```

Now that we got our posterior we can look at some interesting tables and plots

```{r}
plot(m1_gor, rotate_x_labels = T)+
  scale_x_discrete(
  labels = c(
    "Turnip",
    "Cucumber",
    "Carrot",
    "Apple",
    "Grape",
    "Tomato"
  ))
```

Getting a probability table

```{r}
get_probabilities_table(m1_gor, format='html')
```

# Automated labeling

Let's start importing the data

```{r}
d2 <- read_csv('./data/auto-label-comparison.csv') %>%
  dplyr::rename(Value='value') %>% 
  dplyr::filter(ValueType=='accuracy') %>% 
  dplyr::filter(Model=='LabelPropagation_knn' | 
                Model=='LabelPropagation_rbf' |
                Model=='LabelSpreading_knn' |
                Model=='LabelSpreading_rbf')
```

Lets visualize

```{r}
d2 %>% 
  dplyr::sample_n(size=10) %>% 
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "100%")
```

Now we need to convert it to paired comparison: First we pivot it longer and them we expand the dataset

```{r}
d_acc_rank <- d2 %>% 
  dplyr::group_by(Dataset, Variable, Iteration) %>%
  dplyr::mutate(Rank=-rank(Value, ties.method = 'random')) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-Value) %>% #we need to drop the Value variable to pivot wider
  tidyr::pivot_wider(names_from = Model,
                     values_from=Rank)
```

Expanding it

```{r}
models<-unique(d2$Model)
n_models = length(models)
comb <- gtools::combinations(n=n_models, r=2, v=seq(1:n_models), repeats.allowed = F) #all teh paired combinations

d_acc_bt <-  dplyr::tribble(~model0, ~model1, ~y, ~Iteration, ~Dataset, ~DatasetType)

#now we loop each row of the rank wide dataset and create a new one
for(i in 1:nrow(d_acc_rank))
{
  current_row <- d_acc_rank[i,]
  for(j in 1:nrow(comb)){
    comb_row <- comb[j,]
    
    model0 <- models[comb_row[1]]
    model0_rank <- current_row[[1,model0]]
    
    model1 <- models[comb_row[2]]
    model1_rank <- current_row[[1,model1]]
    
    diff_rank <- model1_rank - model0_rank
    #SInce higher accuracy is better if model 1 rank- model 0 rank is positive than model1 wins and y=1 else y=0
    y <- ifelse(diff_rank>0, 1, 0) 
    
    
    d_acc_bt <-d_acc_bt %>%  
      add_row(model0=model0,
              model1=model1,
              y=y,
              Iteration=current_row$Iteration,
              Dataset=current_row$Dataset,
              DatasetType=current_row$DataType)
    
  }
}
```

```{r}
d_acc_bt %>% 
  dplyr::sample_n(size=20) %>% 
  kable() %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "100%")
```

Now we can do some paired comparison models

```{r eval=F}
m1_autol <-
  bpc(
    d_acc_bt,
    player0 = 'model0',
    player1 = 'model1',
    result_column = 'y',
    model_type = 'bt-U',
    cluster = c("Dataset"),
    priors = list(prior_lambda_std = 0.5, prior_U1_std=1.0),
    iter = 3000
  )
save_bpc_model(m1_autol, 'm1_autol', 'fittedmodels')
```

```{r echo=F}
m1_autol <- load_bpc_model('fittedmodels/m1_autol.RDS')
```

```{r}
get_probabilities_table(m1_autol, format = 'html', model_type = 'bt')
```

```{r}
get_rank_of_players_table(m1_autol, format = 'html')
```

# Football

The data comes in the bpcs package

```{r}
data("brasil_soccer_league")
knitr::kable(head(brasil_soccer_league))
```

Let's analyze only the data from 2019 and remove a few columns that are not relevant for this example:

```{r}
d3 <- brasil_soccer_league %>%
  dplyr::filter(Date >= as.Date("2019-01-01") &
                  Date <= as.Date("2019-12-31")) %>%
  dplyr::select(HomeTeam, VisitorTeam, ScoreHomeTeam, ScoreVisitorTeam, Round) 
```

```{r eval=F}
m_football <- bpc(d3, 
          player0 = 'VisitorTeam', 
          player1 =  'HomeTeam', 
          player0_score = 'ScoreVisitorTeam',
          player1_score = 'ScoreHomeTeam',
          model_type = 'davidson',
          solve_ties = 'none',
          priors = list(prior_lambda_std=2.0), # making a more informative prior to improve convergence
          iter = 3000) #stan indicates a low bulk ESS so we are increasing the number of iterations
save_bpc_model(m_football, 'm_football', 'fittedmodels')
```

```{r echo=F}
m_football <- load_bpc_model('fittedmodels/m_football.RDS')
```

Getting the table with the parameters:

```{r}
get_parameters_table(m_football, format = 'html')
```

Ranking the teams

```{r}
get_rank_of_players_table(m_football, format = 'html')
```

# Environment

```{r}
sessionInfo()
```
